Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-04-11 parts/sa.tex
restore procs
fix (again) def of lastprocs
minor fix in lastprocs
remove last minor fix in lastprocs
fixes after Morgan's review

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given $\w\in\OS$, $\concr(\w)$ is the set of minimal scenarios concretizing $\w$ in the
context $\ctx$. It is defined as the largest set satisfying the following conditions:
\begin{enumerate}[(i)]
\item $\forall\delta\in\concr(\w), \exists s\subseteq\ctx, \delta\in\Sce(s)$
\item $\forall\delta\in\concr(\w),
  \exists \phi:\indexes{\w}\to\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m)),$
\\\hspace*{2cm}$\forall n\in\indexes{\w}, \PHbounce(\w_n) \in \ctx\play\delta_{1..\phi(n)}$
\item $\forall\delta,\delta'\in\concr(\w)
				\card{\delta}\leq\card{\delta'} \Rightarrow
					\delta\neq\delta'_{1..\card{\delta}}$.
\end{enumerate}
\end{definition}
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) \DEF
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{lemma}
\label{lem:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{lemma}

@@

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subseteq \Proc \cup \Obj \cup \Sol$ and $\E \subseteq \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a set of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), thus leading to several required processes (\pref{eq:ESol2}).
If $a \in \components$, the reachability of one of its process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a \in \cs$, the reachability of $a_i$ is simply solved by the set of processes $\csState(a_i)$ (\pref{def:csState}) that it represents (\pref{eq:EPrio}).
The solving of an objective $P$ may require a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}); in this case, $P$ is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and~\eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Finally, as the active process of every sort may evolve, $\cwB$ is obtained by iteratively saturating with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
%$$\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \}$$
$\allprocs(\V, \E) = (V \cap \Proc)$.
\todo{Remove $\allprocs$}

@@

In the graph of local causality, an edge $(p, ps) \in \Proc \times \Sol$
is said coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretization of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in \pref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_j$.
\end{definition}

@@

Given an initial context $\ctx$, the reachability of $s$ in $\PH$ is equivalent to the concretization of $\PHobjp{\reach}{0}{1}$ in $\PH'$ from the initial context $\ctx \cup \{ \reach_0 \}$ (the initial state of $\total$ does not matter), which can be efficiently under-approximated using \pref{th:approxinf}.
Indeed, the additional action $\PHhit{\pfp_s(\total)}{\reach_0}{\reach_1}$ in $\PHh'^{(2)}$ allows to conclude on the reachability of process $\pfp_s(\total)$, that is, on the reachability of the state $s$ (considering only the components).

@@

From \pref{th:approxinf}, we can deduce that
for any scenario $\delta$ in $\uconcr(P)$,
there exists a set of processes $ps\in\lastprocs(P)$
such that $ps \subset (s\play\delta)$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
with $\ctx'=\ctx\Cap\procs(\mycwB{\ctx}{P})$,
there exists a scenario $\delta'$ concretizing $\w'$ from the
state $(s\play\delta)$.
Therefore, the scenario $\delta\concat\delta'$ concretizes
$\w$.

\begin{definition}[$\lastprocs$]
\label{def:lastprocs}
Given an objective $\obj{a_i}{a_j}$, $\lastprocs:
\Obj\to\powerset(\powerset(\Proc))$ is
defined such that, $\forall lps\in\lastprocs(\obj{a_i}{a_j})$, 
$lps\in\powerset(\Proc)$,
\begin{enumerate}
\item $a_j\in lps$;
\item $\exists\zeta\in\BS(\obj{a_i}{a_j})$ s.t.
 $\PHsort(\hitter{\zeta_{\card\zeta}}) \neq 
		\PHsort(\bounce{\zeta_{\card\zeta}})
		 \Rightarrow
		  \hitter{\zeta_{\card\zeta}}\in lps$;
\item $\forall b_j\in lps, b\in\cs, \exists ps\in\csState(b_j), ps\subset lps$;
\item $\forall b_j\in lps, b\in\cs, 
		\forall ps\in\csState(b_j),
			\exists lps'\in\lastprocs(\obj{a_i}{a_j}),
				ps\subset lps'$;
\item $\nexists lps'\in\lastprocs(\obj{a_i}{a_j}), lps'\subset lps \wedge
									lps'\neq lps$.
\end{enumerate}
\end{definition}

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a Process Hitting $(\PHs; \PHl; \PHa^{\langle 2 \rangle})$
that satisfies \pref{cr:bounded} and \pref{cr:compcs},
a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in \PHl, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in \PHl, s\subset\ctx$, there exists a scenario
concretizing $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}

>>>>

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given $\w\in\OS$, $\concr(\w)$ is the set of minimal scenarios concretising $\w$ in the
context $\ctx$. It is defined as the largest set satisfying the following conditions:
\begin{enumerate}[(i)]
\item $\forall\delta\in\concr(\w), \exists s\subseteq\ctx, \delta\in\Sce(s)$
\item $\forall\delta\in\concr(\w),
  \exists \phi:\indexes{\w}\to\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m)),$
\\\hspace*{2cm}$\forall n\in\indexes{\w}, \PHbounce(\w_n) \in \ctx\play\delta_{1..\phi(n)}$
\item $\forall\delta,\delta'\in\concr(\w)
				\card{\delta}\leq\card{\delta'} \Rightarrow
					\delta\neq\delta'_{1..\card{\delta}}$.
\end{enumerate}
\end{definition}
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) \DEF
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{lemma}
\label{lem:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{lemma}

@@

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subseteq \Proc \cup \Obj \cup \Sol$ and $\E \subseteq \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a set of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), thus leading to several required processes (\pref{eq:ESol2}).
If $a \in \components$, the reachability of one of its process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a \in \cs$, the reachability of $a_i$ is simply solved by the set of processes $\csState(a_i)$ (\pref{def:csState}) that it represents (\pref{eq:EPrio}).
The solving of an objective $P$ may require a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}); in this case, $P$ is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and~\eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Finally, as the active process of every sort may evolve, $\cwB$ is obtained by iteratively saturating with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
\[\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \} \enspace.\]

@@

In the graph of local causality, an edge $(p, ps) \in \Proc \times \Sol$
is said to be coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretisation of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in \pref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said to be coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_j$.
\end{definition}

@@

Given an initial context $\ctx$, the reachability of $s$ in $\PH$ is equivalent to the concretisation of $\PHobjp{\reach}{0}{1}$ in $\PH'$ from the initial context $\ctx \cup \{ \reach_0 \}$ (the initial state of $\total$ does not matter), which can be efficiently under-approximated using \pref{th:approxinf}.
Indeed, the additional action $\PHhit{\pfp_s(\total)}{\reach_0}{\reach_1}$ in $\PHh'^{(2)}$ allows to conclude on the reachability of process $\pfp_s(\total)$, that is, on the reachability of the state $s$ (considering only the components).

@@

From \pref{th:approxinf}, we can deduce that
for any scenario $\delta$ in $\uconcr(P)$,
there exists a set of processes $ps\in\lastprocs(P)$
such that $ps \subset (s\play\delta)$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
with $\ctx'=\ctx\Cap\procs(\mycwB{\ctx}{P})$,
there exists a scenario $\delta'$ concretising $\w'$ from the
state $(s\play\delta)$.
Therefore, the scenario $\delta\concat\delta'$ concretises
$\w$.

\begin{definition}[$\lastprocs$]
\label{def:lastprocs}
Given an objective $\obj{a_i}{a_j}$, $\lastprocs:
\Obj\to\powerset(\powerset(\Proc))$ is
defined as the largest set such that, $\forall lps\in\lastprocs(\obj{a_i}{a_j})$, 
$lps\in\powerset(\Proc)$,
\begin{enumerate}
\item $a_j\in lps$;
\item $\exists\zeta\in\BS(\obj{a_i}{a_j}),
    \PHsort(\hitter{\zeta_{\card\zeta}}) \neq 
		\PHsort(\bounce{\zeta_{\card\zeta}})
		 \Rightarrow
		  \hitter{\zeta_{\card\zeta}}\in lps$;
\item $\forall b_j\in lps, b\in\cs, \exists ps\in\csState(b_j), ps\subset lps$;
\item $\nexists lps'\in\lastprocs(\obj{a_i}{a_j}), lps'\subset lps \wedge
									lps'\neq lps$.
\end{enumerate}
\end{definition}

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a Process Hitting $(\PHs; \PHl; \PHa^{\langle 2 \rangle})$
that satisfies \pref{cr:bounded} and \pref{cr:compcs},
a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in \PHl, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in \PHl, s\subset\ctx$, there exists a scenario
concretising $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}

### 2014-04-11 parts/suppl-sa.tex
fixes after Morgan's review

  \item Suppose all children objectives of $P$ are concretizable.
    If $\exists (P, Q) \in \Bee{\Obj}{\Obj}$, then by hypothesis,
      $\muconcr_{s}(\obj{\PHtarget(P)}{\PHtarget(Q)} \concat Q) \neq \emptyset$, thus
      $\muconcr_{s}(P) \neq \emptyset$.
    Else, by \pref{def:maxCont}, the concretizations of the children of $P$ require no process of sort $\PHsort(P)$.
      Furthermore, there exists $\zeta \in \BS(P)$ so that $(P, \aZ) \in \Bee{\Obj}{\Sol}$.
      We show by induction that for all $n \in \indexes{\zeta}$, there is a scenario $\delta_n$ so that $\PHget{(s \PHplay \delta_n)}{\PHsort(P)} = \PHbounce(\zeta_n)$.
      \begin{itemize}
        \item[$\circ$] Suppose that $\delta_n$ exists and let $\zeta_n = \PHhit{b_i}{a_j}{a_k}$.
        By hypothesis there exists $\delta' \in \muconcr_{s \PHplay \delta_n}(\PHobj{\any}{b_i})$ with $\PHsort(P) \notin \PHsort(\delta')$ (by \pref{def:maxCont}).
        By \pref{lem:update} there exists $\delta'' \in \restriction{\Sce}{1}(s \PHplay \delta')$ so that $\update(s \PHplay \delta') = s \PHplay \delta' \PHplay \delta''$.
        Furthermore, $\PHget{(s \PHplay \delta' \PHplay \delta'')}{b} = b_j$ (by
		\pref{lem:hcompcomp} if $b \in \components$ or \pref{lem:hcscomp} if $b \in \cs$).
        Therefore, $\delta_{n+1} = \delta_n \PHplay \delta' \PHplay \delta'' \PHplay \zeta_n$.
      \end{itemize}
      Thus, $\delta_{|\zeta|} \in \muconcr_s(P)$. % and $\ceil(\delta) \subseteq max\ctx$.
\end{itemize}
Finally, as $\muconcr_{max\ctx}(\w) \neq \emptyset$, $\uconcr(\w) \neq
\emptyset$ (\pref{lem:uconcr-ctx}).
\end{proofapproxinf}

>>>>

  \item Suppose all children objectives of $P$ are concretisable.
    If $\exists (P, Q) \in \Bee{\Obj}{\Obj}$, then by hypothesis,
      $\muconcr_{s}(\obj{\PHtarget(P)}{\PHtarget(Q)} \concat Q) \neq \emptyset$, thus
      $\muconcr_{s}(P) \neq \emptyset$.
    Else, by \pref{def:maxCont}, the concretisations of the children of $P$ require no process of sort $\PHsort(P)$.
      Furthermore, there exists $\zeta \in \BS(P)$ so that $(P, \aZ) \in \Bee{\Obj}{\Sol}$.
      We show by induction that for all $n \in \indexes{\zeta}$, there is a scenario $\delta_n$ so that $\PHget{(s \PHplay \delta_n)}{\PHsort(P)} = \PHbounce(\zeta_n)$.
      \begin{itemize}
        \item[$\circ$] Suppose that $\delta_n$ exists and let $\zeta_n = \PHhit{b_i}{a_j}{a_k}$.
        By hypothesis there exists $\delta' \in \muconcr_{s \PHplay \delta_n}(\PHobj{\any}{b_i})$ with $\PHsort(P) \notin \PHsort(\delta')$ (by \pref{def:maxCont}).
        By \pref{lem:update} there exists $\delta'' \in \restriction{\Sce}{1}(s \PHplay \delta')$ so that $\update(s \PHplay \delta') = s \PHplay \delta' \PHplay \delta''$.
        Furthermore, $\PHget{(s \PHplay \delta' \PHplay \delta'')}{b} = b_j$ (by
		\pref{lem:hcompcomp} if $b \in \components$ or \pref{lem:hcscomp} if $b \in \cs$).
        Therefore, $\delta_{n+1} = \delta_n \PHplay \delta' \PHplay \delta'' \PHplay \zeta_n$.
      \end{itemize}
      Thus, $\delta_{|\zeta|} \in \muconcr_s(P)$. % and $\ceil(\delta) \subseteq max\ctx$.
\end{itemize}
Finally, as $\muconcr_{max\ctx}(\w) \neq \emptyset$, $\uconcr(\w) \neq
\emptyset$ (\pref{lem:uconcr-ctx}).
\end{proofapproxinf}
